Grappa: Scalable Computing on Mass-Market Hardware for Irregular, Low-Locality Applications

### 2012-07-22 doc/asplos13/evaluation.tex
Minor edits to eval

\subsubsection{Unbalanced tree search in memory}
We ran UTS-mem on Grappa and the XMT with a binomial and two geometric
100M-vertex trees (T3L,T1L,T2L), a geometric 1.6B-vertex tree (T1XL), and a geometric
4.2B-vertex tree (T1XXL).
Both systems do relatively poorly on the binomial tree because it is very deep and a
traversal produces too little parallelism to get good bandwidth.
Figure~\ref{fig:grappa-xmt-uts}
shows the performance in terms of number of vertices visited per
second versus number of compute nodes. The Grappa results shown are for 5
cores per node and the best parameter values from a search over
runtime and application parameters.
Grappa with \checkme{16} machines is faster than the entire XMT. Grappa achieves
\checkme{188Mvert/s} with \checkme{128 machines} and the
XMT achieves only 50Mvert/s, plateauing at 60 nodes. Beyond 90 nodes,
Grappa adds 1.4 Mvert/s/node. The XMT scales at 850 Kvert/s/node,
until it plateaus. Grappa keeps scaling up through 128---the maximum we
could use---although scaling declines because of the unscalability of our aggregation
mechanism as number of network endpoints increases. Using 6 cores per
node gave Grappa up to a 20\% benefit at each data point, but Grappa could
not run well beyond 90 machines with this core configuration.

@@

Despite our efforts to tune the UTS implementation specific to Cray
XMT, performance does not scale well with increasing processor count,
flattening out around sixty processors.  When we increase the size of
the tree from 100M to 4.2B, we find that performance does not improve, suggesting that performance is not limited by task parallelism.   Cray's performance tools show that an increasing number of memory retry operations are generated by code within the runtime system.  Retries are performed by the memory controller when remote synchronization operations fail to find the full-bit associated with each memory location in the unavailable state.  Retries are issued at low priority relative to new memory operations issued by the processor by other contexts, so they consume what would otherwise be unused injection bandwidth.  On a full-bandwidth system such as the MTA-2, retries have no impact on the progress of tasks other than their own.  On a Cray XMT, network bandwidth is limited, so retries create congestion.  In comparison, Grappa performs synchronization without retries, delaying response at the receiving end until ready to notify the sender to proceed.  This saves bandwidth and permits scaling of tasks performing synchronization even on low injection rate networks.

@@

Figure~\ref{fig:bfs-performance} shows Grappa's performance on BFS
compared to that of the XMT\footnote{We do not know the reason for the
  XMT's performance dip at 112 nodes, but it is repeatable}. There are
two things to observe. First, At 128 XMT processors/cluster nodes,
the XMT is  4.17 times faster than Grappa. This is likely due to the
XMT's faster network: the XMT is able to issue 100 million network
requests per second per node; at the 128 nodes datapoint, Grappa is
aggregating 25 million network requests per second, or 4 times slower.
Second, Grappa is scaling as we add processors.

@@

Figure~\ref{fig:centrality-performance} shows Grappa's performance on
Betweenness Centrality compared to that of the XMT. At 128 XMT
processors/cluster nodes, the XMT is 1.75 times faster than Grappa.

\subsection{Benefit of aggregator}

@@

To evaluate the benefit of the aggregator, we ran two
experiments. First, we ran a simple unidirectionaly ping test to see
the maximum benefit the aggregator can provide in terms of improved
network efficiency. Second, we ran BFS with the aggregator disabled in
order to measure its benefit on an application.

@@

First, small message performance against the existing libraries is
terrible. The MPI application test shows us that peak per node
bandwidth supported by our Infiniband card is 3.4GB/s. This is
achievable only with large messages; we must send 16KB packets to get
within 5 percent of peak bandwidth. But in our benchmarks, we saw
average message between 32 and 64 bytes. At 32 bytes, the MPI test is
using less than 7 percent of its peak bandwidth. Grappa sending
messages directly through GASNet uses less than 3 percent of the peak
bandwidth.

Second, aggregation has the potential to improve this situation by an
order of magnitude. With aggregation, Grappa is able to send 32-byte
messags over 12 times faster than using GASNet directly. This is a
more respectable 32 percent of peak bandwidth. Due to expedient design
decisions, Grappa's aggregator limits its aggregation to 4KB; this
limits its peak achievable bandwidth to to 75 percent of the actual
peak.

@@

\paragraph{Aggreator timeout}

>>>>

\subsubsection{Unbalanced tree search in memory}

We ran UTS-mem on Grappa and the XMT with a binomial and two geometric
100M-vertex trees (T3L,T1L,T2L), a geometric 1.6B-vertex tree (T1XL),
and a geometric 4.2B-vertex tree (T1XXL). Both systems do relatively
poorly on the binomial tree because it is very deep and a traversal
produces too little parallelism. Figure~\ref{fig:grappa-xmt-uts} shows
the performance in terms of number of vertices visited per second versus
number of compute nodes. The Grappa results shown are for 5 cores per
node and the best parameter values from a search over runtime and
application parameters. Grappa with \checkme{16} machines is faster than
the entire XMT of 128 processors. Grappa achieves \checkme{188Mvert/s}
with \checkme{128 machines} and the XMT achieves only 50Mvert/s,
plateauing at 60 nodes. Beyond 90 nodes, Grappa adds 1.4 Mvert/s/node.
The XMT scales at 850 Kvert/s/node, until it plateaus. Grappa keeps
scaling up through 128---the maximum we could use---although scaling
declines because of the unscalability of our aggregation mechanism as
number of network endpoints increases. Using 6 cores per node gave
Grappa up to a 20\% benefit at each data point, but Grappa could not run
well beyond 90 machines with this core configuration because of the
current network aggregator design.

@@

Despite our efforts to tune the UTS implementation specific to the 
XMT, performance does not scale well with increasing processor count,
flattening out around sixty processors.  When we increase the size of
the tree from 100M to 4.2B, we find that performance does not improve, suggesting that performance is not limited by task parallelism.   Cray's performance tools show that an increasing number of memory retry operations are generated by code within the runtime system.  Retries are performed by the memory controller when remote synchronization operations fail to find the full-bit associated with each memory location in the unavailable state.  Retries are issued at low priority relative to new memory operations issued by the processor by other contexts, so they consume what would otherwise be unused injection bandwidth.  On a full-bandwidth system such as the MTA-2, retries have no impact on the progress of tasks other than their own.  On a Cray XMT, network bandwidth is limited, so retries create congestion.  In comparison, Grappa performs synchronization without retries, delaying responses at the receiving end until ready to notify the sender to proceed.  This saves bandwidth and permits scaling of tasks performing synchronization even on low injection rate networks.

@@

Figure~\ref{fig:bfs-performance} shows Grappa's performance on BFS
compared to that of the XMT\footnote{The performance dip at 112 nodes is not an anomaly.  It is highly repeatable.  Despite extensive investigation on our part, we can find no reasonable explanation for it.}. There are
two things to observe. First, At 128 XMT processors/cluster nodes,
the XMT is  4.17 times faster than Grappa. This is due to the
XMT's faster network: the XMT is able to issue 100 million network
requests per second per node; at the 128 nodes datapoint, Grappa is
aggregating 25 million network requests per second, or 4 times slower.
\checkme{Second, Grappa is scaling...}

@@

Figure~\ref{fig:centrality-performance} shows Grappa's performance on
Betweenness Centrality compared to that of the XMT. At 128 XMT
processors/cluster nodes, the XMT is 1.75 times faster than Grappa.
\TODO{Add more here...}

\subsection{Network aggregation performance}

@@

To evaluate the benefits of network aggregation, we ran two experiments.
First, we ran a simple unidirectional ping test to see the maximum
benefit the aggregator can provide in terms of improved network
efficiency. Second, we ran BFS with the aggregator disabled in order to
measure its benefit on an application.

@@

First, small message performance against the existing libraries is, as expected, poor. The MPI application test shows us that peak per node
bandwidth supported by our infiniband card is 3.4GB/s. This is
achievable only with large messages; we must send 16KB packets to get
within 5 percent of peak bandwidth. But in our benchmarks, we saw
average message between 32 and 64 bytes. At 32 bytes, the MPI test is
using less than 7 percent of its peak bandwidth. Grappa sending
messages directly through GASNet uses less than 3 percent of the peak
bandwidth.

Second, aggregation has the potential to improve this situation by an
order of magnitude. With aggregation, Grappa is able to send 32-byte
messages over 12 times faster than using GASNet directly. This is a
more respectable 32 percent of peak bandwidth. Due to expedient design
decisions, Grappa's aggregator limits its aggregation to 4KB; this
limits its peak achievable bandwidth to 75 percent of the actual
peak.

@@

\paragraph{Aggregator timeout}
